---
  title: "Exercises - Wrangling Data at scale with data.table"
output: html_document
date: "`r format(Sys.time(), '%d %B %Y')`"
html_document:
  toc: TRUE
df_print: paged
number_sections: FALSE
highlight: tango
theme: lumen
toc_depth: 3
toc_float: true
css: custom.css 
self_contained: false
includes:
  after_body: footer.html
---
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.align = 'center', echo = TRUE)
```

***

```{r}
library(data.table)
library(coronavirus)
library(dplyr)
```

# Now, we will use data.table and the coronavirus package and compare it with the dplyr package.

# 1a - Transform a coronavirus dataset into a data.table. Name it 'corona'. Now, return the 10 last row. 
```{r}

```

# 1b How would you do that but with a data frame and using dplyr?
```{r}

```

# 2a Filter the table into one that has information about European countries. Name it Euro_corona. Subset all rows from the data.table where the country is Germany.
```{r}

```

# 2b How would you subset using a data frame?
```{r}

```

# 3a Use %like% to search for Gibraltar. 
```{r}

```

# 3b How would you do that in a data frame?
```{r}

```

# 4a Use %between% to check all rows between June and July
```{r}

```

# 4b How would you do that in a data frame?
```{r}

```

# 5a Now use %chin%. Subset all rows where country is Norway, Denmark, and Sweden.
```{r}

```

# 5b How would you do that with a data frame?
```{r}

```

# 6a Select the first three columns. Don't forget to check the class.
```{r}

```

# 6b How would you deselect them?
```{r}

```
