---
  title: "Exercises - Wrangling Data at scale with data.table"
output: html_document
date: "`r format(Sys.time(), '%d %B %Y')`"
html_document:
  toc: TRUE
df_print: paged
number_sections: FALSE
highlight: tango
theme: lumen
toc_depth: 3
toc_float: true
css: custom.css 
self_contained: false
includes:
  after_body: footer.html
---
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.align = 'center', echo = TRUE)
```

***

```{r}
library(data.table)
library(coronavirus)
library(dplyr)
```

# Now, we will use data.table and the coronavirus package and compare it with the dplyr package.

# 1a - Transform a coronavirus dataset into a data.table. Name it 'corona_dt'. Now, return the 10 last rows. 
```{r}
data("coronavirus")

```

# 1b What's another way to do this using other tools we know?
```{r}

```

# 2a Filter the table into one that has information about European countries. Name it euro_corona. Subset all rows from the data.table where the country is Germany.
```{r}

```

# 2b How would you subset using dplyr functions?
```{r}

```

# 3a Use %like% to return all provinces that include "island" in the name. Can you report how many provinces that includes?
```{r}

```

# 3b How would you do the above with dplyr::filter()?
```{r}

```

# 4a Now try using %chin%. Subset all rows in euro_corona where country is Norway, Denmark, and Sweden.
```{r}

```

# 5a Select the first three columns. Don't forget to check the class.
```{r}

```

# 5b How would you deselect them?
```{r}

```

# 5c Create a new column called pop_size that indicates whether a country population is "large" (over 10 million) or "small" (below 10 million). Don't worry too much about inclusive/exclusive on the numbers here.
```{r}

```

# 6. Bringing it all together: Let's say you want to create a new column that reports the a daily incidence rate, which is calculated thus: incidence_rate = (confirmed_cases / population) * 100,000

# How would you go about that?
```{r}

```


